Optimal Fitting of Planar Curves to Prescribed Constraints

### 2014-02-02 idp/documents/documentation/paper.tex
fixed some spelling errors

% TODO: expand n't and *'s contractions
% TODO: check if all quotations are done `like this'

@@

		When trying to improve something, it seems advisable to take a step back and thoroughly research the causes of its perceived shortcomings so as to find a way to improve it without making the same or similar mistakes. In the context of curve design tools, this can be achieved by analyzing and modelling the process that people use when designing curves. Once such a model has been obtained, it can be used to propose criteria for good curve design tools as well as to analyze usability aspects of existing tools.

@@

			Step 1 is of course independent of any software that may be used by the designer. Step 2 and 3 are strongly influenced by the choice of design tool, in that the software dictates which properties it accepts as specifications for curves (the specification language the software uses). Unless the properties the designer conveyed to the software in step 3 uniquely identify the curve, in step 4, the design tool will have to choose the curve that the user was most likely describing from a possibly infinite set of curves that exhibit the supplied properties. This notion of `most likely' is usually modelled in terms of a fairness measure, fairer curves being regarded as more likely. Fairness ususally encodes intuitive notions such as smoothness and minimality using mathematical concepts such as continuity and variation of curvature. Here, the software has another opportunity to influence the design process both positively and negatively depending on what fairness measure is chosen and how good the software is at reliably finding a curve that has a high fairness.

			Note that there may be multiple ways to describe a given curve design tool in terms of this process. Instead of modelling the selection of the most likely curve using some fairness measure, one may also propose an alternate view on the process in which the user implicitly specifies the exact curve they want in the properties they specify. For example, in a curve design tool which takes two points from the user and connects them using a straight line, one may say that the user provides the start and the end point of the curve, and the fairness measure is designed to always select the straight line as the fairest curve. Another way of looking at this process is to say that the user, by supplying \(p_0\) and \(p_1\), specified the coefficients of the linear term \(p_0 + \xp{p_1 - p_0} t\), thus uniquely identifying the desired curve. Modelling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool.

@@

				When applying the model introduced in section \ref{section:curve_design_process} to Bézier spline design tools, there are two possibilities for modelling the fairness measure. One way to view things is to assume that the user, in supplying the four coefficient points of each curve segment, uniquely specifies each segment's Bézier curve and thus the entire spline, leaving no room for choosing a curve according to some fairness measure. Since users usually don't think about Bézier splines that way, it is more practical to assume that the user only supplies both point and velocity at the start and end of each curve segment, with a fairness measure designed to always choose the cubic Bézier curve that is uniquely specified by the given properties as the fairest curve.

@@

				The description language used to specify Bézier splines in common design tools is lacking in expressiveness. Specifically, describing curves in terms of their curvature is difficult since curvature may only be specified indirectly through the magnitude of the velocity values (higher velocity resulting in lower curvature). Unfortunately, this is also very dependent on the point and velocity coefficients at the other end of the spline segment. There is also no way to directly specify constant or linearly changing curvature. This makes specifying curvature very tedious, and in case the winding angle of the curve segment is large enough, actually impossible, forcing the designer to use additional nodes in order to get the desired result. Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created or moved in order to change the overall curve. It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes. Another issue is that without software support for curvature continuity, using more nodes usually also introduces more points where curvature is not continuous. In cases where the curvature is an important property of the curve (for example the constant curvature of circular arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points.

@@

				Again, the model from section \ref{section:curve_design_process} is applied. With Spiro splines, the modelling is fairly straightforward. The user supplies a set of points for which the software computes an interpolating spline. The fairness measure chooses the segments of this spline to be parts of the Euler spiral and makes sure that there is curvature continuity and no excessive winding.

				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_criteria_curve_design_tools}. The specification language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them are needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3). The fairness measure leads to beautifully smooth curves that are ususally sufficiently close to what a human would expect given the control points (criteria 4 and 5). The software is also good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 6 well enough. The only real issue lies in the insufficient expressiveness of the description language (criterion 1), discussed further in the following paragraph.

				Section \ref{section:bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversively affects usability. Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify direction nor curvature. These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature. This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itsself excellently to the expression of circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature. Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties. Unfortunately, this is rarely practical, since it both requires a lot of time to set up and doesn't behave very well when modified. These issues are somewhat alleviated by the guaranteed curvature continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved. That way, when indirectly specifying tangent angle and/or curvature by placing many control points, it's not as easy to accidently create small bumps as with Bézier splines.

@@

				Let \(n\) be the number of specification items. Let \(\hat{t}_i\) be the position of the \(i\)th specification item. If the \(i\)th item is a point specification, the point is denoted \(\hat{\phi}_i\), if it is a direction specification, the direction is denoted \(\hat{\delta}_i\) and if it is a curvature specification, the curvature is doneted \(\hat{\chi}_i\). The specified total arc length for the curve is denoted \(\hat{\lambda}\).

@@

				Once the optimization problem is put together, it is passed to the opimization solver, which will return the optimized coefficients of the segment polynomials. These segment polynomials are then assembled according to the layout described in section \ref{section:segmentation_notation} in order to obtain the final curve.

@@

				The optimization process is perfomed in a number of steps. Depending on the change of the specification, only certain steps have to be run again as previous results can be reused.

@@

				\subsubsection{Optimization Problem}
					Next, the optimization problem is created. The optimization problem specifies an objective function which includes the speed error discussed in section \ref{section:constant_speed} as well as the fairness function discussed in section \ref{section:fairness_measure}. Furthermore the continuity connections discussed in section \ref{section:continuity_connections} and the satisfaction of the specifications of the curve are realized using a constraint function. This step also computes the first and second derivatives of those functions. The concrete values required by the curve specifications are not used; instead, placeholder variables are introduced. For the total length of the curve a placeholder is used as well. This allows the reuse of the optimization problem if curve specifications or curve length change; the derivatives do not have to be recomputed. Thus, the optimization problem is cached and only rebuilt if the optimiztion segments change or if the number of each type of constraint on the curve changes. If the value of a constraint changes, but the type of constraint remains the same (e.g. a point constraint requires the curve to pass through a different point than before), the optimization problem does not have to be rebuilt.

@@

					As Kurve uses a background thread to optimize curves, it is not possible to use the Curve objects yieded by the optimizer to render the curves in the main thread at the same time. For that reason, Curve objects are converted to DiscreteCurve objects in that background process.

@@

				The user interface does not focus on user friendliness. Its purpose is to test the validity of the approach to curve design researched in this project. For that reason, most actions are performed using keyboard shortcuts. Many user interface elments can be selected with the mouse; this enables to use keyboard shortcuts that affect that component only. It is possible to interact with some components using the mouse. However, there is no toolbar or similar user interface elements that would make any of the features more accessible.

@@

					To save all curves in the window, press `S'. A save file dialog will appear that allows the user to specify a filename. Saved files can later be loaded by pressing `L' which displays an open file dialog. Keep in mind that loading a file dismisses all curves in the current window.

@@

				% TODO: make this smaller, maybe use landscape format
				\begin{figure}[p]
					\centering
					\includegraphics[width=\textwidth]{content/output/ui_speeds.pdf}
					\caption{Example of a Curve that is `Overspecified' (i.e. the speed of the curve cannot be kept constant due to the high amount of constraints)}
					\label{figure:ui_speeds}
				\end{figure}

				\subsubsection{Command Cheat Sheet}
					\begin{itemize}
						\item General
						\begin{description}
							\item[`N'] Add a new curve
							\item[`L'/`S'] Load/Save
							\item[`1'/`2'] decrease/increase internal segment count
							\item[Alt + `1'/`2'] decrease/increase polynomial template degree
						\end{description}
						\item All Controls
						\begin{description}
							\item[Click] Selects a control 
							\item[Shift + Click] Select multiple controls
							\item[Shift + Scroll (+ Alt)] Insert length at component (slowly)
						\end{description}
						\item Specification Control
						\begin{description}
							\item[Ctrl + Scroll (+ Alt)] Change the position of the specification (slowly)
							\item[Shift + Drag] Change position of specification (adjusts the point of the specification)
							\item[`R'] Removes the specification
							\item[`P'] Triggers point specification
							\item[Drag (+ Alt)] Move point (slowly). Only possible when point specification is enabled.
							\item[`D'] Triggers direction specification
							\item[Scroll (+ Alt)] Changes the direction (slowly). Only possible when direction specification is enabled.
							\item[`C'] Triggers curvature specification
							\item[Win + Scroll (+ Alt)] Changes the curvature (slowly). Only possible when curvature specification is enabled.
							\item[Shift + Click] Select multiple controls
							\item[Shift + Scroll (+ Alt)] Insert length at component (slowly)
						\end{description}
						\item Segment Control
						\begin{description}
							\item[Right-Click] Add empty specifcation control at that position
							\item[Drag] Moves any point specifications delimiting the segment. (Only possible if at leats one of the specification controls has the point specification enabled.)
						\end{description}
					\end{itemize}

@@

				The ability to specify curvature constitues a significant improvement over common curve design tools, where this is usually not possible. Using this functionality, the user can specify curves which are best expressed in terms of curvature in a natural and concise way. Not only does this make the curve design process less tedious, but it may also lead to better curves, since only the essential information about the curve is given to the software. This allows the fairness measure to choose the best curve, something that is not possible if lots of auxiliary information is supplied to the software in order to indirectly influence the curvature.

@@

				Section \ref{section:specification_language} describes why it makes sense to specify the curve's total arc length. While fixing the arc length of the curve (and thus also the arc length between each specification item) may seem a little restrictive at first, there are actually quite a few upsides. For instance, by adjusting the arc length of a curve with a point specification item on each end, one can conveniently produce curcular arcs between two points, since increasing the arc length will make the curve bulge out, with the fairness making sure that a circular arc is produced. Generally speaking, having control over the arc length of the curve results in benign and predictable curve behavior and is often useful for pulling it in place, not unlike the way one would handle a mechanical spline.

@@

				As a final way of demonstrating the high expressiveness of the description language, it may be noted that it actually allows for the description of arbitrary segments of the Euler spiral. Since every segment of the Euler spiral exhibits linearly changing curvature between its start and end points, specififying these curvature values on an otherwise unconstrained curve will cause the fairness measure to select the desired segment of the Euler spiral.

@@

				Another issue is reliability. For some descriptions, the iterative optimization algorithm may converge extremely slowly or not at all, thus not producing a curve. In some cases, it is possible that small changes in the description restore the ability to converge on a solution, but sometimes one simply has to start over. There are also cases where the optimization gets stuck in a pathological local optimum, in which case one often has to start over aswell.

@@

			\begin{figure}[htb]
				\centering
				\begin{subfigure}[b]{8 \textwidth / 13}
					\includegraphics[width=\textwidth]{content/output/examples_prototype_fluttershy.pdf}
					\caption{Fluttershy \cite{fluttershy}}
					\label{figure:examples_prototype_fluttershy}
				\end{subfigure}%
				\begin{subfigure}[b]{5 \textwidth / 13}
					\includegraphics[width=\textwidth]{content/output/examples_prototype_fairness.pdf}
					\caption{Fairness Measure Example}
					\label{figure:examples_prototype_fairness}
				\end{subfigure}
				\caption{Examples of curves designed using the prototype curve design tool. Point specification items are displayed as red dots, direction specification items are displayed green tangent lines and curvature specification items are displayed using blue circular arcs taken from the osculating circle, or tangent lines if the curve is locally a straght line. The thin blue lines visualize the curvature at each point of the curve (longer lines meaning higher curvature).}
				\label{figure:examples_prototype}
			\end{figure}

			Figure \ref{figure:examples_prototype_fluttershy} shows a few curves designed using the prototype curve design tool. Example 1 demonstrates how adjusting the arc length of an otherwise unconstrained curve allows the convenient creation of curcular arcs between two points.

@@

			It may also be worth estimating the amount of information given in the description. Recall that the curve in example 4 was already designed using Bézier splines in section \ref{section:bézier_splines}, using 7 nodes and 12 handles, and Spiro splines in section \ref{section:spiro_splines}, using 7 control points. Here, it is designed using 2 point specifications, 1 direction specification, and 3 curvature specifications and specified arc length. Unfortunately, it is somewhat difficult to directly compare these amounts since specification items use fractions of the arc length for positioning instead of node/control point ordering, as is the case with Bézier splines and Spiro splines. Fraction of arc length positions do of course contain more information than simple orderings, although this is arguably not the case if said positions lie exactly on the start or end of the curve. Furthermore, direction and curvature specification items only specify a scalar value instead of a plane vector, so could be considered to only supply half as much information as a point specification item. So we count 1 scalar value for arc length, 4 scalar values for points, 1 for directions and 1 for curvatures specified at the start or end of the curve, and 4 scalar values for curvatures specified in the middle of the curve, giving us a total of 11 scalar values for the entire specification, just shy of the 14 scalar values for the Spiro spline and a lot less than the 33 scalar values for the Bézier spline (where the loss of one degree of freedom for smooth nodes with colinear velocities has been taken into account). Of course these comparisons cannot be taken as precise, quantitative measurements of conciseness, but they are useful as a coarse judgement on the amount of information that needs to be supplied to define a specific curve.

@@

			It follows immediately that every parametric curve that is \(\contp{n}\) continuous is also \(\contg{n}\) continuous. One may also observe that \(\contp{0}\) continuity and \(\contg{0}\) continuity are equivalent, since discontinuities in the parametrization function itsself manifest themselves directly in the curve it represents, and thus have to exist in all parametrizations of this curve. In the case of parametric plane curves, using the definitions from section \ref{section:parametric_plane_curves}, we can equate \(\contg{0}\) continuity with \(\phi\) being continuous, \(\contg{1}\) continuity with \(\delta\) being continuous and \(\contg{2}\) continuity with \(\chi\) being continuous.

>>>>

@@

		When trying to improve something, it seems advisable to take a step back and thoroughly research the causes of its perceived shortcomings so as to find a way to improve it without making the same or similar mistakes. In the context of curve design tools, this can be achieved by analyzing and modeling the process that people use when designing curves. Once such a model has been obtained, it can be used to propose criteria for good curve design tools as well as to analyze usability aspects of existing tools.

@@

			Step 1 is of course independent of any software that may be used by the designer. Step 2 and 3 are strongly influenced by the choice of design tool, in that the software dictates which properties it accepts as specifications for curves (the specification language the software uses). Unless the properties the designer conveyed to the software in step 3 uniquely identify the curve, in step 4, the design tool will have to choose the curve that the user was most likely describing from a possibly infinite set of curves that exhibit the supplied properties. This notion of `most likely' is usually modeled in terms of a fairness measure, fairer curves being regarded as more likely. Fairness usually encodes intuitive notions such as smoothness and minimality using mathematical concepts such as continuity and variation of curvature. Here, the software has another opportunity to influence the design process both positively and negatively depending on what fairness measure is chosen and how good the software is at reliably finding a curve that has a high fairness.

			Note that there may be multiple ways to describe a given curve design tool in terms of this process. Instead of modeling the selection of the most likely curve using some fairness measure, one may also propose an alternate view on the process in which the user implicitly specifies the exact curve they want in the properties they specify. For example, in a curve design tool which takes two points from the user and connects them using a straight line, one may say that the user provides the start and the end point of the curve, and the fairness measure is designed to always select the straight line as the fairest curve. Another way of looking at this process is to say that the user, by supplying \(p_0\) and \(p_1\), specified the coefficients of the linear term \(p_0 + \xp{p_1 - p_0} t\), thus uniquely identifying the desired curve. Modeling the process in such a way that the user supplies only the minimum amount of implicit information, relying on the fairness measure for everything else, is usually closer to reality when analyzing usability of some curve design tool.

@@

				When applying the model introduced in section \ref{section:curve_design_process} to Bézier spline design tools, there are two possibilities for modeling the fairness measure. One way to view things is to assume that the user, in supplying the four coefficient points of each curve segment, uniquely specifies each segment's Bézier curve and thus the entire spline, leaving no room for choosing a curve according to some fairness measure. Since users usually don't think about Bézier splines that way, it is more practical to assume that the user only supplies both point and velocity at the start and end of each curve segment, with a fairness measure designed to always choose the cubic Bézier curve that is uniquely specified by the given properties as the fairest curve.

@@

				The description language used to specify Bézier splines in common design tools is lacking in expressiveness. Specifically, describing curves in terms of their curvature is difficult since curvature may only be specified indirectly through the magnitude of the velocity values (higher velocity resulting in lower curvature). Unfortunately, this is also very dependent on the point and velocity coefficients at the other end of the spline segment. There is also no way to directly specify constant or linearly changing curvature. This makes specifying curvature very tedious, and in case the winding angle of the curve segment is large enough, actually impossible, forcing the designer to use additional nodes in order to get the desired result. Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created or moved in order to change the overall curve. It also sacrifices smoothness, since it's easy to accidentally introduce subtle bumps in the curve when using a large number of nodes. Another issue is that without software support for curvature continuity, using more nodes usually also introduces more points where curvature is not continuous. In cases where the curvature is an important property of the curve (for example the constant curvature of circular arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points.

@@

				Again, the model from section \ref{section:curve_design_process} is applied. With Spiro splines, the modeling is fairly straightforward. The user supplies a set of points for which the software computes an interpolating spline. The fairness measure chooses the segments of this spline to be parts of the Euler spiral and makes sure that there is curvature continuity and no excessive winding.

				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_criteria_curve_design_tools}. The specification language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them are needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3). The fairness measure leads to beautifully smooth curves that are usually sufficiently close to what a human would expect given the control points (criteria 4 and 5). The software is also good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 6 well enough. The only real issue lies in the insufficient expressiveness of the description language (criterion 1), discussed further in the following paragraph.

				Section \ref{section:bézier_splines} discussed how the lack of ability to specify curvature using common Bézier spline design tools adversely affects usability. Many of the same points apply to Spiro spline design tools, since they allow the user to neither specify direction nor curvature. These properties can only be expressed indirectly, by placing the control points in such a way that the algorithm will create a curve that has the desired properties, resulting in the usability issues described previously, such as requiring many control points to be specified even though the curve could be described much more concisely in terms of direction and/or curvature. This seems especially unfortunate when considering that the spline primitive of Spiro splines, the Euler spiral, would lend itself excellently to the expression of circular arcs, spirals and other shapes exhibiting constant or linearly changing curvature. Sometimes, the algorithm can be coaxed into producing the desired curve by placing just a few control points very carefully, resulting in a series of Euler spiral segments that exhibit the wanted properties. Unfortunately, this is rarely practical, since it both requires a lot of time to set up and doesn't behave very well when modified. These issues are somewhat alleviated by the guaranteed curvature continuity together with the general tendency of the algorithm to produce extremely smooth curves even when many control points are involved. That way, when indirectly specifying tangent angle and/or curvature by placing many control points, it's not as easy to accidentally create small bumps as with Bézier splines.

@@

				Let \(n\) be the number of specification items. Let \(\hat{t}_i\) be the position of the \(i\)th specification item. If the \(i\)th item is a point specification, the point is denoted \(\hat{\phi}_i\), if it is a direction specification, the direction is denoted \(\hat{\delta}_i\) and if it is a curvature specification, the curvature is denoted \(\hat{\chi}_i\). The specified total arc length for the curve is denoted \(\hat{\lambda}\).

@@

				Once the optimization problem is put together, it is passed to the optimization solver, which will return the optimized coefficients of the segment polynomials. These segment polynomials are then assembled according to the layout described in section \ref{section:segmentation_notation} in order to obtain the final curve.

@@

				The optimization process is performed in a number of steps. Depending on the change of the specification, only certain steps have to be run again as previous results can be reused.

@@

				\subsubsection{Optimization Problem}
					Next, the optimization problem is created. The optimization problem specifies an objective function which includes the speed error discussed in section \ref{section:constant_speed} as well as the fairness function discussed in section \ref{section:fairness_measure}. Furthermore the continuity connections discussed in section \ref{section:continuity_connections} and the satisfaction of the specifications of the curve are realized using a constraint function. This step also computes the first and second derivatives of those functions. The concrete values required by the curve specifications are not used; instead, placeholder variables are introduced. For the total length of the curve a placeholder is used as well. This allows the reuse of the optimization problem if curve specifications or curve length change; the derivatives do not have to be recomputed. Thus, the optimization problem is cached and only rebuilt if the optimization segments change or if the number of each type of constraint on the curve changes. If the value of a constraint changes, but the type of constraint remains the same (e.g. a point constraint requires the curve to pass through a different point than before), the optimization problem does not have to be rebuilt.

@@

					As Kurve uses a background thread to optimize curves, it is not possible to use the Curve objects yielded by the optimizer to render the curves in the main thread at the same time. For that reason, Curve objects are converted to DiscreteCurve objects in that background process.

@@

				The user interface does not focus on user friendliness. Its purpose is to test the validity of the approach to curve design researched in this project. For that reason, most actions are performed using keyboard shortcuts. Many user interface elements can be selected with the mouse; this enables to use keyboard shortcuts that affect that component only. It is possible to interact with some components using the mouse. However, there is no toolbar or similar user interface elements that would make any of the features more accessible.

@@

					To save all curves in the window, press `S'. A save file dialog will appear that allows the user to specify a file name. Saved files can later be loaded by pressing `L' which displays an open file dialog. Keep in mind that loading a file dismisses all curves in the current window.

@@

				% TODO: make this smaller, maybe use landscape format
				\begin{figure}[p]
					\centering
					\includegraphics[width=\textwidth]{content/output/ui_speeds.pdf}
					\caption{Example of a Curve that is `overspecified' (i.e. the speed of the curve cannot be kept constant due to the high amount of constraints)}
					\label{figure:ui_speeds}
				\end{figure}

				\subsubsection{Command Cheat Sheet}
					\begin{itemize}
						\item General
						\begin{description}
							\item[`N'] Add a new curve
							\item[`L'/`S'] Load/Save
							\item[`1'/`2'] decrease/increase internal segment count
							\item[Alt + `1'/`2'] decrease/increase polynomial template degree
						\end{description}
						\item All Controls
						\begin{description}
							\item[Click] Selects a control 
							\item[Shift + Click] Select multiple controls
							\item[Shift + Scroll (+ Alt)] Insert length at component (slowly)
						\end{description}
						\item Specification Control
						\begin{description}
							\item[Ctrl + Scroll (+ Alt)] Change the position of the specification (slowly)
							\item[Shift + Drag] Change position of specification (adjusts the point of the specification)
							\item[`R'] Removes the specification
							\item[`P'] Triggers point specification
							\item[Drag (+ Alt)] Move point (slowly). Only possible when point specification is enabled.
							\item[`D'] Triggers direction specification
							\item[Scroll (+ Alt)] Changes the direction (slowly). Only possible when direction specification is enabled.
							\item[`C'] Triggers curvature specification
							\item[Win + Scroll (+ Alt)] Changes the curvature (slowly). Only possible when curvature specification is enabled.
							\item[Shift + Click] Select multiple controls
							\item[Shift + Scroll (+ Alt)] Insert length at component (slowly)
						\end{description}
						\item Segment Control
						\begin{description}
							\item[Right-Click] Add empty specification control at that position
							\item[Drag] Moves any point specifications delimiting the segment. (Only possible if at least one of the specification controls has the point specification enabled.)
						\end{description}
					\end{itemize}

@@

				The ability to specify curvature constitutes a significant improvement over common curve design tools, where this is usually not possible. Using this functionality, the user can specify curves which are best expressed in terms of curvature in a natural and concise way. Not only does this make the curve design process less tedious, but it may also lead to better curves, since only the essential information about the curve is given to the software. This allows the fairness measure to choose the best curve, something that is not possible if lots of auxiliary information is supplied to the software in order to indirectly influence the curvature.

@@

				Section \ref{section:specification_language} describes why it makes sense to specify the curve's total arc length. While fixing the arc length of the curve (and thus also the arc length between each specification item) may seem a little restrictive at first, there are actually quite a few upsides. For instance, by adjusting the arc length of a curve with a point specification item on each end, one can conveniently produce circular arcs between two points, since increasing the arc length will make the curve bulge out, with the fairness making sure that a circular arc is produced. Generally speaking, having control over the arc length of the curve results in benign and predictable curve behavior and is often useful for pulling it in place, not unlike the way one would handle a mechanical spline.

@@

				As a final way of demonstrating the high expressiveness of the description language, it may be noted that it actually allows for the description of arbitrary segments of the Euler spiral. Since every segment of the Euler spiral exhibits linearly changing curvature between its start and end points, specifying these curvature values on an otherwise unconstrained curve will cause the fairness measure to select the desired segment of the Euler spiral.

@@

				Another issue is reliability. For some descriptions, the iterative optimization algorithm may converge extremely slowly or not at all, thus not producing a curve. In some cases, it is possible that small changes in the description restore the ability to converge on a solution, but sometimes one simply has to start over. There are also cases where the optimization gets stuck in a pathological local optimum, in which case one often has to start over as well.

@@

			\begin{figure}[htb]
				\centering
				\begin{subfigure}[b]{8 \textwidth / 13}
					\includegraphics[width=\textwidth]{content/output/examples_prototype_fluttershy.pdf}
					\caption{Fluttershy \cite{fluttershy}}
					\label{figure:examples_prototype_fluttershy}
				\end{subfigure}%
				\begin{subfigure}[b]{5 \textwidth / 13}
					\includegraphics[width=\textwidth]{content/output/examples_prototype_fairness.pdf}
					\caption{Fairness Measure Example}
					\label{figure:examples_prototype_fairness}
				\end{subfigure}
				\caption{Examples of curves designed using the prototype curve design tool. Point specification items are displayed as red dots, direction specification items are displayed green tangent lines and curvature specification items are displayed using blue circular arcs taken from the osculating circle, or tangent lines if the curve is locally a straight line. The thin blue lines visualize the curvature at each point of the curve (longer lines meaning higher curvature).}
				\label{figure:examples_prototype}
			\end{figure}

			Figure \ref{figure:examples_prototype_fluttershy} shows a few curves designed using the prototype curve design tool. Example 1 demonstrates how adjusting the arc length of an otherwise unconstrained curve allows the convenient creation of circular arcs between two points.

@@

			It may also be worth estimating the amount of information given in the description. Recall that the curve in example 4 was already designed using Bézier splines in section \ref{section:bézier_splines}, using 7 nodes and 12 handles, and Spiro splines in section \ref{section:spiro_splines}, using 7 control points. Here, it is designed using 2 point specifications, 1 direction specification, and 3 curvature specifications and specified arc length. Unfortunately, it is somewhat difficult to directly compare these amounts since specification items use fractions of the arc length for positioning instead of node/control point ordering, as is the case with Bézier splines and Spiro splines. Fraction of arc length positions do of course contain more information than simple orderings, although this is arguably not the case if said positions lie exactly on the start or end of the curve. Furthermore, direction and curvature specification items only specify a scalar value instead of a plane vector, so could be considered to only supply half as much information as a point specification item. So we count 1 scalar value for arc length, 4 scalar values for points, 1 for directions and 1 for curvatures specified at the start or end of the curve, and 4 scalar values for curvatures specified in the middle of the curve, giving us a total of 11 scalar values for the entire specification, just shy of the 14 scalar values for the Spiro spline and a lot less than the 33 scalar values for the Bézier spline (where the loss of one degree of freedom for smooth nodes with colinear velocities has been taken into account). Of course these comparisons cannot be taken as precise, quantitative measurements of conciseness, but they are useful as a coarse judgment on the amount of information that needs to be supplied to define a specific curve.

@@

			It follows immediately that every parametric curve that is \(\contp{n}\) continuous is also \(\contg{n}\) continuous. One may also observe that \(\contp{0}\) continuity and \(\contg{0}\) continuity are equivalent, since discontinuities in the parametrization function itself manifest themselves directly in the curve it represents, and thus have to exist in all parametrizations of this curve. In the case of parametric plane curves, using the definitions from section \ref{section:parametric_plane_curves}, we can equate \(\contg{0}\) continuity with \(\phi\) being continuous, \(\contg{1}\) continuity with \(\delta\) being continuous and \(\contg{2}\) continuity with \(\chi\) being continuous.
